A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-03-11 survey.tex
Minor typos

A typical StarCraft map is defined as a rectangular grid, where the $width \times height$ of the map is measured in the number of $32 \times 32$ square of pixels, a.k.a build tiles. Although the
resolution  at which  units  move  is in squares of $8 \times 8$ pixels, a.k.a walk tiles. The typical dimensions for maps range from $64 \times 64$ to $256 \times 256$ build tiles. Each player  can
control up  to 200 units (not including  buildings).  Moreover, each
different race  contains between 30  to 35 different types  of units
and  buildings,  each  one  with  a significant  number  of  special
abilities. All these factors  together make StarCraft a significant
challenge,   in   which   humans    are   still   much   better   than
computers. For instance, in the game ladder iCCup\footnote{http://www.iccup.com/StarCraft/} where users are ranked by their current point totals ($E$ being the lowest possible rank, and $A^+$ and $Olympic$ being the second highest and highest ranks, respectively), the best StarCraft AI bots are ranked between $D$ and $D^+$, where average  amateur players are  ranked between $C^+$ and  $B$. For comparison, StarCraft professional  players are usually ranked between
$A^-$ and $A^+$.

From a theoretical point of view, the state space of a StarCraft for a given map is enormous. For example, consider a 128x128 map. At any given moment there might be between 50 to 400 units in the map, each of which might have a complex internal state (remaining energy and hit-points, action being executed, etc.). This quickly leads to an immense number of possible states (way beyond the size of smaller games, such as Chess or Go). For example, just considering the location of each unit (with 128x128 possible positions per unit), and 400 units, gives us an initial number of $16384^{400} \approx 10^{1685}$. If we add the other factors playing a role in the game, we obtain even larger numbers.

@@

\subsubsection{Task Decomposition}
For all the previous reasons, most existing approaches to play games like StarCraft work by decomposing the problem of playing a RTS game into a collection of smaller problems, to be solved independently. Specifically, a common subdivision is:
\begin{itemize}
\item {\em Strategy}: corresponds to the high-level decision making process.  This is
  the  highest level of  abstraction for  the game  comprehension. Finding   an   efficient   strategy   or
  counter-strategy  against a  given opponent is key in RTS games.  It concerns  the whole
  set of  units and buildings  a player owns. %If  we may allow  ourselves a  military analogy, strategy is orders given by generals at the HQ. 
\item  {\em Tactics}: are  the implementation  of the  current  strategy. It
  implies  army and  building positioning,  movements, timing,  and so
  on.  Tactics concerns a  group of  units. %They  are orders  given by officers on the battlefield.
\item  {\em Reactive  control}: is  the  implementation  of  tactics.  This
  consists   in  moving,   targeting,  firing,   fleeing,  hit-and-run
  techniques (a.k.a. ``kiting'')  during battle. Reactive control focuses
  on a specific unit. %It is soldiers applying officers' orders. 
%\end{itemize}
%
%\noindent
%Both     strategy      and     tactics     are      usually     called
%\textit{macro-management},  whereas reactive  control  is also  called
%\textit{micro-management}. In addition of the three tasks shown above,
%one could also consider these two tasks that an AI has to deal with:
%\begin{itemize}
\item {\em Terrain analysis}: consists  in the analysis of regions composing
  the  map:  choke-points,  minerals  and  gas emplacements,  low  and  high
  walkable grounds, islands, etc.
\item {\em Intelligence gathering}:  corresponds to information collected about
  the  opponent. Because of  the fog-of-war,  players must
  regularly send scouts to localize and spy enemy bases.%, and must see or  infer what buildings and technology the enemy has developed in order  to adapt its strategy.
\end{itemize}

@@

% (Santi: I commented this one, since it's mentioned above) On influence maps, \cite{teamCompositionRTS} studied team composition and maneuvering by learning a self-organizing map, while \cite{HagelbackJ08} presented a multi-agent potential field based bot.

@@

% POTENTIAL FIELDS:
Potential fields (or influence maps) have been found to be a useful technique for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multi-agent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

@@

Started in 2010, the AIIDE StarCraft AI Competition\footnote{\url{http://www.StarCraftAICompetition.com}} is the most
well known and longest running StarCraft AI Competition in the world. Each year,
AI bots are submitted by competitors to do battle within the retail version of
StarCraft: BroodWar, with prizes supplied by Blizzard Entertainment. %This competition has been made possible by the BWAPI StarCraft programming interface, which allows players to write programs which can retrieve game data and issue game commands to StarCraft via an easy to use C++ API.

@@

Tournament 4 was the complete game of {\em StarCraft: BroodWar} with fog-of-war enforced. The tournament
was run with a random pairing double-elimination format with each match being best of 5 games.
Competitors could play as any of the three races, with the only limitations in gameplay
being those which were considered ``cheating'' in the StarCraft community. A map pool of
5 well-known professional maps was announced to competitors in advance, with a random map being chosen for each game.

@@

An initial attempt to run a StarCraft tournament at the Computational
Intelligence in Games conference (CIG 2010) suffered from technical problems.
These mainly stemmed from the desire to use evolved, largely untested
maps which proved to look interesting but made the submitted bots and the Broodwar Terrain Analyzer (BWTA) provided with the BWAPI interface crash
so frequently that it would have been unjustifiable to announce a winner.

@@

These maps came in two flavors, namely three 3-player maps:
\emph{Athena-II}, \emph{Neo Moon Glaive}, \emph{Tears of the Moon},
and three 6-player maps:
\emph{Legacy}, \emph{River of Light}, and 
\emph{The Huntress 1.1}. 
We shall note that some bots consistently crashed on one
of the originally considered maps which has thus been replaced.
This is surprising as all maps are well known league play maps
or have been provided with the StarCraft Broodwar distribution
itself.
Setup, replays and results for the CIG 2012 competition can be found
here\footnote{\url{http://ls11-www.cs.tu-dortmund.de/rts-competition/StarCraft-cig2012}}.

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder1.pdf}
    \caption{StarCraft Bot Ladder: bot's Elo Rating from February 1, 2012 to June 20, 2012}
    \label{fig:botLadder1}
\end{figure*}

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder2.pdf}
    \caption{StarCraft Bot Ladder: bot's Elo Rating from September 1, 2012 to January 10, 2013}
    \label{fig:botLadder2}
\end{figure*}

>>>>

A typical StarCraft map is defined as a rectangular grid, where the $width \times height$ of the map is measured in the number of $32 \times 32$ square of pixels, also knows as build tiles. Although the
resolution  at which  units  move  is in squares of $8 \times 8$ pixels, also knows as walk tiles. The typical dimensions for maps range from $64 \times 64$ to $256 \times 256$ build tiles. Each player  can
control up  to 200 units (not including  buildings).  Moreover, each
different race  contains between 30  to 35 different types  of units
and  buildings,  each  one  with  a significant  number  of  special
abilities. All these factors  together make StarCraft a significant
challenge,   in   which   humans    are   still   much   better   than
computers. For instance, in the game ladder iCCup\footnote{http://www.iccup.com/StarCraft/} where users are ranked by their current point totals ($E$ being the lowest possible rank, and $A^+$ and $Olympic$ being the second highest and highest ranks, respectively), the best StarCraft AI bots are ranked between $D$ and $D^+$, where average  amateur players are  ranked between $C^+$ and  $B$. For comparison, StarCraft professional  players are usually ranked between
$A^-$ and $A^+$.

From a theoretical point of view, the state space of a StarCraft for a given map is enormous. For example, consider a $128 \times 128$ map. At any given moment there might be between 50 to 400 units in the map, each of which might have a complex internal state (remaining energy and hit-points, action being executed, etc.). This quickly leads to an immense number of possible states (way beyond the size of smaller games, such as Chess or Go). For example, just considering the location of each unit (with $128 \times 128$ possible positions per unit), and 400 units, gives us an initial number of $16384^{400} \approx 10^{1685}$. If we add the other factors playing a role in the game, we obtain even larger numbers.

@@

\subsubsection{Task Decomposition}
For all the previous reasons, most existing approaches to play games like StarCraft work by decomposing the problem of playing a RTS game into a collection of smaller problems, to be solved independently. Specifically, a common subdivision is:
\begin{itemize}
\item {\em Strategy}: corresponds to the high-level decision making process.  This is
  the  highest level of  abstraction for  the game  comprehension. Finding   an   efficient   strategy   or
  counter-strategy  against a  given opponent is key in RTS games.  It concerns  the whole
  set of  units and buildings  a player owns. %If  we may allow  ourselves a  military analogy, strategy is orders given by generals at the HQ. 
\item  {\em Tactics}: are  the implementation  of the  current  strategy. It
  implies  army and  building positioning,  movements, timing,  and so
  on.  Tactics concerns a  group of  units. %They  are orders  given by officers on the battlefield.
\item  {\em Reactive  control}: is  the  implementation  of  tactics.  This
  consists   in  moving,   targeting,  firing,   fleeing,  hit-and-run
  techniques (also knows as ``kiting'')  during battle. Reactive control focuses
  on a specific unit. %It is soldiers applying officers' orders. 
%\end{itemize}
%
%\noindent
%Both     strategy      and     tactics     are      usually     called
%\textit{macro-management},  whereas reactive  control  is also  called
%\textit{micro-management}. In addition of the three tasks shown above,
%one could also consider these two tasks that an AI has to deal with:
%\begin{itemize}
\item {\em Terrain analysis}: consists  in the analysis of regions composing
  the  map:  choke-points,  minerals  and  gas emplacements,  low  and  high
  walkable grounds, islands, etc.
\item {\em Intelligence gathering}:  corresponds to information collected about
  the  opponent. Because of  the fog-of-war,  players must
  regularly send scouts to localize and spy enemy bases.%, and must see or  infer what buildings and technology the enemy has developed in order  to adapt its strategy.
\end{itemize}

@@

% (Santi: I commented this one, since it's mentioned above) On influence maps, \cite{teamCompositionRTS} studied team composition and maneuvering by learning a self-organizing map, while \cite{HagelbackJ08} presented a multiagent potential field based bot.

@@

% POTENTIAL FIELDS:
Potential fields (or influence maps) have been found to be a useful technique for reactive decision making. Some uses of potential fields in RTS games are: avoiding obstacles (navigation), avoiding opponent fire \cite{uriarte2012kiting}, or staying at maximum shooting distance \cite{Hagelback09}. Potential fields have also been combined with A* path-finding to avoid local traps \cite{Hagelback12}. Hagelb\"{a}ck and Johansson \cite{HagelbackJ08} presented a multiagent potential fields based bot able to deal with fog-of-war in the Tankbattle game. Avery et al. \cite{Avery09} and Smith et al. \cite{SmithCIG10} co-evolved influence map trees for spatial reasoning in RTS games. Danielsiek et al. \cite{Danielsiek_2008} used influence maps to achieve intelligent squad movement to flank the opponent in a RTS game. Despite their success, a drawback for potential field-based techniques is the large number of parameters that has to be tuned in order to achieve the desired behavior. Approaches for automatically learning such parameters have been explored, for example, using reinforcement learning \cite{Liu_2008}, or self-organizing-maps (SOM) \cite{teamCompositionRTS}. We would like to note that potential fields are a reactive control technique, and as such, they do not perform any form of lookahead. As a consequence, these techniques are prone to make units stuck in local maxima.

@@

Started in 2010, the AIIDE StarCraft AI Competition\footnote{\url{http://www.StarCraftAICompetition.com}} is the most
well known and longest running StarCraft AI Competition in the world. Each year,
AI bots are submitted by competitors to do battle within the retail version of
StarCraft: Brood War, with prizes supplied by Blizzard Entertainment. %This competition has been made possible by the BWAPI StarCraft programming interface, which allows players to write programs which can retrieve game data and issue game commands to StarCraft via an easy to use C++ API.

@@

Tournament 4 was the complete game of {\em StarCraft: Brood War} with fog-of-war enforced. The tournament
was run with a random pairing double-elimination format with each match being best of 5 games.
Competitors could play as any of the three races, with the only limitations in gameplay
being those which were considered ``cheating'' in the StarCraft community. A map pool of
5 well-known professional maps was announced to competitors in advance, with a random map being chosen for each game.

@@

An initial attempt to run a StarCraft tournament at the Computational
Intelligence in Games conference (CIG 2010) suffered from technical problems.
These mainly stemmed from the desire to use evolved, largely untested
maps which proved to look interesting but made the submitted bots and the Brood War Terrain Analyzer (BWTA) provided with the BWAPI interface crash
so frequently that it would have been unjustifiable to announce a winner.

@@

These maps came in two flavors, namely three 3-player maps:
\emph{Athena-II}, \emph{Neo Moon Glaive}, \emph{Tears of the Moon},
and three 6-player maps:
\emph{Legacy}, \emph{River of Light}, and 
\emph{The Huntress 1.1}. 
We shall note that some bots consistently crashed on one
of the originally considered maps which has thus been replaced.
This is surprising as all maps are well known league play maps
or have been provided with the StarCraft Brood War distribution
itself.
Setup, replays and results for the CIG 2012 competition can be found
here\footnote{\url{http://ls11-www.cs.tu-dortmund.de/rts-competition/StarCraft-cig2012}}.

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder1.pdf}
    \caption{StarCraft Bot Ladder: bot's ELO Rating from February 1, 2012 to June 20, 2012}
    \label{fig:botLadder1}
\end{figure*}

\begin{figure*}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/botLadder2.pdf}
    \caption{StarCraft Bot Ladder: bot's ELO Rating from September 1, 2012 to January 10, 2013}
    \label{fig:botLadder2}
\end{figure*}
